---
title: "Strategic Market Analysis"
author: "Karen, Samuel, Laurence"
date: "February 8, 2019"
output: 
        html_document:
        keep_md: yes
--- 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- install packages hidden-->
```{r echo=FALSE, include=FALSE}

    install.packages('dplyr', repos="http://cran.us.r-project.org")
    install.packages('knitr', repos="http://cran.us.r-project.org")
    install.packages('rvest', repos="http://cran.us.r-project.org")
    install.packages('ggplot2', repos="http://cran.us.r-project.org")
    install.packages('psych', repos="http://cran.us.r-project.org")
    install.packages('maps', repos="http://cran.us.r-project.org")
    install.packages('plyr', repos="http://cran.us.r-project.org")
    
    library(dplyr)
    library(knitr)
    library(rvest)
    library(ggplot2)
    library(psych)
    library(maps)
    library(plyr)

```
## Introduction
As of the latest 2015 data, AB InBev enjoys a 45.8% Share of the US beer market. While the merger with InBev has made Anheuser-Busch the largest brewer in America, when aggregate data is considered AB InBev has in fact seen a total reduction of 1.4% in its share of the US beer market when compared to 2014 data, this despite the fact that worldwide, AB InBev has seen a steady growth in market share. One factor for this may be the growing micro brew movement and the influx of craft beers into the market. AB InBev's market strategy has been and continues to be the marketing and support of its core Brands and segment growth through acquisition of established brands. However, an analysis of market competition seems to indicates that the micro brew movement has grown from a novelty product to a serious brand competitor. This analysis will explore AB InBev's emerging competition by analyzing the breweries and beers currently competing for US consumers.  


## The Data
Data was derived from a nationwide dataset of 558 US breweries and a matching dataset of 2,410 US craft beers. We analyzed the data and compared it to state geographic data to determine the distribution of the breweries by state. We then analyzed the alcohol by volume (ABV) and international bitterness units (IBU) of the sampled beers to identify trends, characteristics,  correlations.  


```{r ReadFile}

#Read in the beers.csv and set to data frame
Beers <- read.csv("C:/CaseStudy1/Beers.csv", header = TRUE, sep = ",")

#Read in the Breweries.csv and set to data frame
Breweries <- read.csv("C:/CaseStudy1/Breweries.csv", header = TRUE, sep = ",")

#Read in the states.csv and set to data frame
States <- read.csv("C:/CaseStudy1/states.csv", header = TRUE, sep = ",")
```
  
## Analysis Tools for R-Studio

To conduct the research we used R-studio as our primary statistical analysis software and loaded the following statistical packages:
      
  library(dplyr)  
  library(knitr)  
  library(rvest)  
  library(ggplot2)  
  library(psych)  
  library(maps)  
  library(plyr)  
  
  
<!-- cleaning up the data hidden-->
```{r merging, echo=FALSE, include=FALSE}
#add equivalent name to Beers data frame for merge
Beers$Brew_ID <- Beers$Brewery_id

#merge the data frames
dataset_merge <- merge(Breweries,Beers, by="Brew_ID",all=TRUE)

#clean up dataset_merge
BeerData <- dataset_merge
BeerData$Brewery_id <- NULL
BeerData$Brew_ID <- NULL
BeerData$Beer_ID <- NULL
colnames(BeerData) <- c('Brewery','City','State','Beer','ABV','IBU','Style','Ounces')
BeerData <- BeerData[c(1,2,3,4,7,5,6,8)]
BeerData <- BeerData[order(BeerData$State), ]
rownames(BeerData) <- NULL

#Create new CSV of merged data
write.csv(BeerData, file = "BeerData.csv")
```

  
## From Alaska to Wyoming A Revolution is Brewing
Nationwide there has been a tremendous increase in the number of microbreweries that have entered the market over the last five years. According to data from the Brewer's Association the number of microbreweries in the US jumped from 620 in 2010 to 2,626 in 2015. 

###### *https://www.brewersassociation.org/statistics/number-of-breweries/
```{r}

#reduce data set to Brewery, State, and Beer
beers <- BeerData[c(1,3,5)]

#pull and filter sample beers and display from head
Alaska <- head(beers,10)
P1 <- Alaska[2:7, ]

#pull and filter sample beers and display from tail
Wyoming <- tail(beers)
P2 <- Wyoming[1:6, ]

knitr::kable(list(P1, P2), row.names = FALSE)
```

## Data Driven Brewing
Thoug small, most of the breweries are making complex beers for an ever growing market of consumers that want to know more about beer. All 2,410 beers in the data set had a specified style and type, 97% of the beers had a specified ABV, and 58% of the beers had a specified IBU.      
```{r}
colSums(!is.na(BeerData[4:7]))
```


## Breweries per state

<!-- create beerMap hidden code-->
```{r BeerMap, echo=FALSE, include=TRUE}

    library("maps")

map.text("state", regions=c("alabama", "arizona", "arkansas", "california", "colorado", "connecticut", "delaware", "florida", "georgia", "idaho", "illinois", "indiana", "iowa", "kansas", "kentucky", "louisiana", "maine", "maryland", "massachusetts:main", "michigan:north", "minnesota", "mississippi", "missouri", "montana", "nebraska", "nevada", "new hampshire", "new jersey", "new mexico", "new york:main", "north carolina:main", "north dakota", "ohio", "oklahoma", "oregon", "pennsylvania", "rhode island", "south carolina", "south dakota", "tennessee", "texas", "utah", "vermont", "virginia:main", "washington:main", "west virginia", "wisconsin", "wyoming"), labels=as.character(c(10, 47, 5, 183, 265, 27, 2, 58, 16, 30, 91, 139, 30, 23, 21, 19, 27, 21, 82, 162, 55, 11, 42, 40, 25, 11, 8, 8, 14, 74, 59, 3, 49, 19, 125, 100, 27, 14, 7, 6, 130, 26, 27, 40, 68, 2, 87, 15)))


```

<!-- clean up brewery by state data hidden-->
```{r BreweriesPerState, echo=FALSE, include=TRUE}
#find number of breweries in state
#library(plyr)
#library(maps)
BreweryState <- count(dataset_merge,"State")
BreweryState <- BreweryState[-c(1,8,12), ]
BreweryState <- BreweryState[order(BreweryState$State), ]
rownames(BreweryState) <- NULL
States <- States[-c(2,11), ]
States <- States[order(States$State), ]
rownames(States) <- NULL
BreweryState$StateLong <- States$StateLong
BrewStateLst <- BreweryState[c(3,2)]
BrewStateLst <- BrewStateLst[order(BrewStateLst$StateLong), ]
colnames(BrewStateLst) <- c("State", "Breweries")

#BrewStateLst2 <- BrewStateLst[order(BrewStateLst$Breweries, decreasing=TRUE), ]
#print(BrewStateLst2[1:10, ], row.names=FALSE) 
#knitr::kable(BrewStateLst2[1:10, ], row.names = FALSE)

L1 <- BrewStateLst[1:16, ]
L2 <- BrewStateLst[17:32, ]
L3 <- BrewStateLst[33:48, ]

knitr::kable(list(L1, L2, L3), row.names = FALSE)

```

## Breweries Per State Graph

```{r statesize}
#Plot of Breweries per State
library(ggplot2)
b1 <- ggplot(BreweryState, aes(x=reorder(BreweryState$State,-BreweryState$freq), y=BreweryState$freq, fill=BreweryState$State)) +
  
geom_bar(stat="identity") + labs(title ="Breweries in States", x= "States", y="Number of Breweries", fill="States") + theme(text = element_text(size=7),axis.text.x = element_text(angle=90, hjust=1))
b1
```


## Compute the median alcohol content and international bitterness unit for each state.


```{r medians}
#Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare. 
library(psych)
ABV_St <- describeBy(dataset_merge$ABV, dataset_merge$State, mat=T)
IBU_St <- describeBy(dataset_merge$IBU, dataset_merge$State, mat=T)
head(ABV_St)
head(IBU_St)
```
## Barplots of median alcohol by volume and IBU  

```{r Final}
#bar plot of Median Alcohol by volumne
ABV_plot <- ggplot(ABV_St, aes(x=reorder(ABV_St$group,-ABV_St$median), y=ABV_St$median, fill=ABV_St$group1)) +
  geom_bar(stat="identity") + labs(title ="Median Alcohol by Volume in States", x= "States", y="ABV", fill="States") + theme(text = element_text(size=9),axis.text.x = element_text(angle=90, hjust=1))
ABV_plot
#bar plot of Median IBU in states
IBU_plot <- ggplot(IBU_St, aes(x=reorder(IBU_St$group,-IBU_St$median), y=IBU_St$median, fill=IBU_St$group1)) +
  geom_bar(stat="identity") + labs(title ="Median Intl Bitterness in States", x= "States", y="IBU", fill="States") + theme(text = element_text(size=9),axis.text.x = element_text(angle=90, hjust=1))	
IBU_plot
#Kentucky KY and DC show to be hightes  median ABV
summary(ABV_St)
#max IBU shown to be Maine (ME)
summary(IBU_St)
```

## Final Plots

```{r finalplots}
library(ggplot2)
#merge by group1 (states) for plot
ABV_IBU_merge <- merge(ABV_St,IBU_St, by="group1",all=TRUE)
#scatter plot of ABV and IBU
med_scatter <- ggplot(ABV_IBU_merge, aes(x=ABV_IBU_merge$median.x, y=ABV_IBU_merge$median.y)) + geom_point(color="blue") + labs(title="Median Alcohol by Volume vs International Bitterness",x="Median Alcohol by Volume", y = "Median Intl Bitterness")+   theme_classic()  
med_scatter 
#scatter w/ regression line
med_scat_reg <- ggplot(ABV_IBU_merge, aes(x=ABV_IBU_merge$median.x, y=ABV_IBU_merge$median.y)) + geom_point(color="blue") + labs(title="Median Alcohol by Volume vs International Bitterness",x="Median Alcohol by Volume", y = "Median Intl Bitterness")+   theme_classic() + stat_ellipse() + geom_smooth(method=lm)
med_scat_reg 
#point bplot by state
med_plot_by_state <- ggplot(ABV_IBU_merge, aes(x = ABV_IBU_merge$group)) + 
  geom_point(aes(y = ABV_IBU_merge$median.x *1000), colour="blue") + 
  geom_point(aes(y = ABV_IBU_merge$median.y), colour = "red") + 
  ylab(label="Alcohol by volume and Intl Bitterness") + 
  xlab("States") + theme(text = element_text(size=9),axis.text.x = element_text(angle=90, hjust=1)) + ggtitle("Median Alcohol by volume(Blue) x 1000 and International Bitterness(Red) by State")
med_plot_by_state
#R standard for comparison
plot(dataset_merge$ABV, dataset_merge$IBU, main="Alcohol by Vol vs Bitterness", xlab="ABV", ylab="IBU")
#summary stats on ABV
summary(dataset_merge$ABV)
summary(ABV_St)
```